### 2013-09-16 paper/kbanotebook2013/content/abstract.tex
Added the skeleton files needed for the trec notebook submission

\begin{abstract}

% TODO this can be updated at the end

In this paper we will present the system design and algorithm adopted by the 
GatorDSR team, University of Florida to efficiently process TREC KBA 2013 --- 
SSF track. Here we will describe the system evolution as well as the details 
of our algorithm to extract slot values for the given slot name. Scalability 
and efficiency in precision and recall are the major   goals of this work, 
given the overly limited time limitation and available computational resources.

\end{abstract}