### 2013-09-16 paper/kbanotebook2013/content/conclusion.tex

\section{Conclusions}

% Note: Give a summary of our efforts and the
% high level contribution to society.

% Talk about the things we wanted to do and
% didnt get to do.