### 2013-09-16 paper/kbanotebook2013/content/discussion.tex

\section{Discussion}

% Note: Here we discuss why the results are the way they are
% Give pros and cons, talk about how the implemented algorithms
% performed in the actual implementation.
% Answer the `why' question about all the trends in the results section.