### 2013-09-16 paper/kbanotebook2013/paper.tex

\documentclass{article}

\begin{document}

\title{GatorDSR: University of Florida Knowledge Base Acceleration Notebook}

\numberofauthors{3}

\author {

\alignauthor Morteza Shahriari\\
  \affaddr{University of Florida}\\
       \affaddr{Department of Computer \& Information Science \& Engineering}\\
       \affaddr{Gainesville, Florida, USA}\\
       \email{\{m.shahriarinia@ufl.edu}

\alignauthor Christan Grant, Yang Peng, Daisy Zhe Wang\\
  \affaddr{University of Florida}\\
       \affaddr{Department of Computer \& Information Science \& Engineering}\\
       \affaddr{Gainesville, Florida, USA}\\
       \email{\{cgrant,ypeng,daisyw\}@cise.ufl.edu}

\alignauthor Milenko Petrovic\\
  \affaddr{Institute for Human and Machine Cognition}\\
       %\affaddr{Department of Computer \& Information Science \& Engineering}\\
       \affaddr{Ocala, Florida, USA}\\
       \email{mpetrovic@ihmc.us}
}

\maketitle

\input{content/abstract.tex}

\input{content/introduction.tex}

\input{content/system.tex}

\input{content/implementation.tex}

\input{content/results.tex}

\input{content/discussion.tex}

\input{content/conclusion.tex}

\section*{Acknowledgements}
Christan Grant is funded by a National Science Foundation Graduate Research
Fellowship under Grant No. DGE-0802270.

\end{document}