### 2013-09-17 paper/kbanotebook2013/content/implementation.tex
Initial copy of the project report in Latex.

\section{Implementation}

We extract aliases for entities from wikipedia automatically both using API and using the actual page content, then apply pattern matching rules for slot value extraction.

\subsection{Alias Generation}

We use Wikipedia API to get some aliases automatically. This is done by retrieving backlink references (redirects of a wiki entity). Unfortunately this is not good enough and to enhance recall we need more aliases. To have better use of a wiki page we parse HTML DOM of the page, then use regular expressions to extract the bold phrases of the first paragraph as alias of the actual entity. Based on our observation this is a very accurate heuristic and provides us with lots of famous aliases of the entities. To consider other typical cases we consider some generic first name last name order swapping conventions such as Bill Gates -> Gates, Bill.

\subsection{Pattern Matching}

In this system, we use pattern matching methods to find corresponding slot values for entities. The procedure is: 1) Find stream items that contain the entities by using alias names of the entities. 2) Inside these stream items, fetch the sentences which contain entities by using alias names and the coreference information provided by Lingpipe tags. 3) Use these sentences to match existing patterns. 4) When patterns matched, generate SSF results.

The format of the patterns is consist of five fields: 1) the type of the entity. 2) the slot name. 3) the pattern content. 4) the direction of the slot valule to the entity. 5) the type of the slot values. When we match one pattern, we match all the fields except the third field. The type of slot values could be the entity type tagged by Linpipe, noun phrase tagged by OpenNLP and the time phrases we hard-code. For these three kinds of patterns, we implement them in different ways accordingly.

Next we will explain the patterns with more details. Lets’s first look at one example of the first kind of patterns. For slot FounderOf, we have this pattern: <PER, FounderOf, founder, right, ORG>. PER means that the entity we are finding slot values for is a PER entity; FounderOf means this is a pattern for FounderOf slot. founder is the ankor word we are trying to match in the sentence text; right means that we are going to the right part of the sentence to match the pattern and find the slot value; ORG means the slot value should be a ORG entity.

Here is an example for the second type: <PER, AwardsWon, awarded, right, NP>, NP meaning noun phrase. This pattern can be interpreted as that we are looking for a noun phrase after the “awarded” since that noun phrase could possibly represent an award. Since titles and awards are usually not the Lingpipe entities, we use the OpenNLP noun phrase chunker to fetch the noun phrases.

Another example for the third type: <PER, DateOfDeath, died, right, last night>. In this pattern, “last night” means we are looking for exactly the phrase “last night” to the right of “died”. This pattern is inspired by the intuition that in news articles, people often mention that somebody died last night instead of mentioning the accurate date information and Lingpipe tends not to tag phrases like “last night” as a DATE entity.

To improve recall, we introduce generic patterns that can generate many results. But these patterns generate many errors, too. Thus there is a trade-off between the accuracy and recall. This is one drawback of using pattern matching method, meaning it is really hard to find good patterns with both high accuracy and recall.

For accuracy, we have found that there are three major kinds of errors: 1) wrong entities found; 2) wrong tags by the Lingpipe; 3) wrong results matched by the patterns. We solve the first problem by using the wikipedia or twitter information of the entities to get better alias names. And we use post-processing to reduce the second and third types of errors.

\subsection{Post Processing}

The SSF output of many extractions is noisy. The data contains duplicates and incorrect extractions. We can define rules to sanitize the output only using the information present in the SSF file. The file is processed in time order, in a tuple-at-a-time fashion to minimize the impact on accuracy. . We define two classes of rules deduplication rules and inference rules.

The output contains many duplicate entries. As we read the list of extrated slots we create rules to define "duplicate". Duplicates can be present in a window of rows; we use a window size of 2 meaning we only be adjacent rows. Two rows are duplicates if 1) they have the same exact extraction, 2) if the rows have the same slot name and a similar slot value and 3) if the extracted sentence for a particular slot types come from the same sentence.

 New slots can be deduced from existing slots by defineing inference rules. For example, two slots for the task are "FounderOf" and "FoundedBy". A safe assumtion is these slot names are biconditional logical connectives with the entities and slot values. Therefore, we can express a rule "X FounderOf Y" equals "Y FoundedBy X" where X and Y are single unique entities. Additionally, we found that the slot names "Contact\_Meet\_PlaceTime" could be infered as "Contact\_Meet\_Entity" if the Entity was a FAC and the extracted sentence contained an additional ORG/FAC tag.

We also remove erronious slots that have extractions that are several pages in length or tool small. Errors of extracting long sentences can typically be attributed to poor sentence parsing of web documents. We have some valid "small" extractions. For example a comma may separate a name and a title (e.g. "John, Professor at MIT"). But such extraction rules can be particularly noisy, so we check to see if the extracted values have good entity values.

The aim in post processing was to sanitize the results increasing the precision of the results by eliminating spurrious results.

\subsection{Shortcomings}

The missing twitter entities are more than the 50\% of the twitter entities we need to work on and this is a very very bad performance. One reason could be that those entities are not very famous. For example Brenda Weiler google search result is 860,000 documents our of billions of web documents. For our small portion of the web it might make sense. If you pay attention to the histogram of the entities found you'll note that more than half of the entities have appeared in less than 10 StreamItems. A good portion have appeared only once. Which the document does not necessarily have good information for us.

A theory is that we are falling behind because we are using the cleansed version of the corpus. We believe there has been a reason that TREC has included the actual HTML document as well as the cleansed version. This definitely will convey some information to us. If it was as easy as reading some clean text they wouldn't bother including so much data for teams to be useless. So we guess is that we are missing some information from not using the actual document. And, we are looking for tokens with entity value set which will depend us dramatically on the accuracy of lingpipe, which is a fast algorithm but is not as good as other NLP tools can be e.g. Stanford NLP.

% Note: Talk about how each of the algorithms were created

% Note: Give enough information for our algorithms to be
% reimplemented and verified.