### 2013-09-17 paper/kbanotebook2013/content/results.tex

\section{Results}

Document extraction using query entity matching with aliases, sentence extraction using alias matching and co-reference. Slot extraction using patterns, NER tags and NP tags. 158,052 documents with query entities, 17885 unique extracted slot values for 8 slots and 139 entities, 4 slots and 31 entities missing.

On the performance of the first submission run we performed some random sampling via two processes, the results of which are as follows:

\begin{table}[b]
\caption{SSF Performance Measure on Initial Submission }
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
 & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Sampling \#1 & 55\% & 17\% & 27\% \\ 
\hline Sampling \#2 & 54\% & 15\% & 31\%  \\ 
\hline 
\end{tabular} 
\end{table}

\begin{table}[b]
\caption{SSF Performance Measure on Submission Infer }
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
 & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Affiliate & \% & \% & \% \\ \hline 
AssociateOf & \% & \% & \%  \\ \hline 
AwardsWon & \% & \% & \%  \\ \hline 
CauseOfDeath & \% & \% & \%  \\ \hline 
Contact\_Meet\_Entity & \% & \% & \%  \\ \hline 
Contact\_Meet\_PlaceTime & \% & \% & \%  \\ \hline 
DateOfDeath & \% & \% & \%  \\ \hline 
EmployeeOf & \% & \% & \%  \\ \hline 
FoundedBy & \% & \% & \%  \\ \hline 
FounderOf & \% & \% & \%  \\ \hline 
Titles & \% & \% & \%  \\ \hline 
TopMembers & \% & \% & \%  \\ \hline

\end{tabular} 
\end{table}

% Note: Here is where we give as much stats as possible on our runs