### 2013-09-18 paper/kbanotebook2013/paper.tex

\documentclass{article}

\usepackage{cite }

\begin{document}

\title{GatorDSR: University of Florida Knowledge Base Acceleration Notebook}

%\numberofauthors{3}

\author {%
  Morteza Shahriari Nia\\
  University of Florida\\
  Department of Electrical \& Computer Engineering\\
  Gainesville, Florida, USA\\
       {m.shahriarinia@ufl.edu}
\and
  Christan Grant, Yang Peng, Daisy Zhe Wang\\
  University of Florida\\
       Department of Computer \& Information Science \& Engineering\\
       Gainesville, Florida, USA\\
       {\{cgrant,ypeng,daisyw\}@cise.ufl.edu}
\and
  Milenko Petrovic\\
  Institute for Human and Machine Cognition\\
       %\affaddr{Department of Computer \& Information Science \& Engineering}\\
       Ocala, Florida, USA\\
       {mpetrovic@ihmc.us}
}

\maketitle

\input{content/abstract.tex}

\input{content/introduction.tex}

\input{content/system.tex}

\input{content/implementation.tex}

\input{content/results.tex}

\input{content/discussion.tex}

\input{content/conclusion.tex}

\section*{Acknowledgements}
Christan Grant is funded by a National Science Foundation Graduate Research
Fellowship under Grant No. DGE-0802270.

%\bibliographystyle{acm}
%\bibliography{citation}

\bibliographystyle{acm}
\bibliography{citation}

\end{document}