### 2013-09-22 paper/kbanotebook2013/content/conclusion.tex

\section{Conclusions}

We experimented through different tools and approaches to best process the massive amounts of data on the platform that we had available to us. We generate aliases for wikipedia entities using Wiki API and extract some aliases from wikipedia pages text itself. On twitter entities we extract aliases manually as it is part of the rule of the KBA track. We process   documents that mention entities for slot value extraction. Slot values are determined using pattern matching over coreferences of entities in sentences. Finally post processing will filter, cleanup and infers some new slot values to enhance recall and accuracy.

We noticed that some tools that claim to be performant for using the hardware capabilities at hand sometimes don't really work as claimed and you should not always rely on one without a thorough A/B testing of performance which we ended up in generating our in-house system for processing the corpus and generating the index. Furthermore, on extracting slot values, pattern matching might not be the best options but definitely can produce some good results at hand. We have plans on generating classifiers for slot value extraction purposes. Entity resolution on the other hand was a topic we spent sometime on but could not get to stable grounds for it. Entity resolution will distinguish between entities of the same name but different contexts. Further improvements on this component of the system are required.

% Note: Give a summary of our efforts and the
% high level contribution to society.

% Talk about the things we wanted to do and
% didnt get to do.