### 2013-09-22 paper/kbanotebook2013/content/results.tex

\section{Results}

Document extraction using query entity matching with aliases, sentence extraction using alias matching and co-reference. Slot extraction using patterns, NER tags and NP tags. 158,052 documents with query entities, 17885 unique extracted slot values for 8 slots and 139 entities, 4 slots and 31 entities missing.

On the performance of the first submission run we performed some random sampling via two processes, the results of which are as follows:

\begin{table}[b]
\caption{SSF Performance Measure on Initial Submission }
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
 & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Sampling \#1 & 55\% & 17\% & 27\% \\ 
\hline Sampling \#2 & 54\% & 15\% & 31\%  \\ 
\hline 
\end{tabular} 
\end{table}

\begin{table}
\caption{SSF Accuracy Measure on Submission Infer }
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
 & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Affiliate & 1\% & 95\% & 5\% \\ \hline 
AssociateOf & 63.6\% & 9.1\% & 27.3\%  \\ \hline 
AwardsWon & 10\% & 10\% & 80\%  \\ \hline 
CauseOfDeath & 0\% & 0\% & 0\%  \\ \hline 
Contact\_Meet\_Entity & 21\% & 42\% & 37\%  \\ \hline 
Contact\_Meet\_PlaceTime & 5\% & 20\% & 85\%  \\ \hline 
DateOfDeath & 29.6\% & 71\% & 25\%  \\ \hline 
EmployeeOf & 5\% & 30\% & 65\%  \\ \hline 
FoundedBy & 62\% & 17\% & 21\%  \\ \hline 
FounderOf & 50\% & 0\% & 50\%  \\ \hline 
Titles & 55\% & 0\% & 45\%  \\ \hline 
TopMembers & 33\% & 17\% & 50\%  \\ \hline

\end{tabular} 
\end{table}

\begin{table}
\caption{Recall Measure on Submission Infer }
\centering
\begin{tabular}{|c|p{3cm}|p{4cm}|}
\hline 
 & Total instances of slot value found & \# of entities covered by slot value \\ 
\hline 
Affiliate & 108598 & 80 \\ \hline 
AssociateOf & 25278 & 106 \\ \hline 
AwardsWon & 38 & 14 \\ \hline 
CauseOfDeath & 0 & 0 \\ \hline 
Contact\_Meet\_Entity & 191 & 8 \\ \hline 
Contact\_Meet\_PlaceTime & 5974 & 109 \\ \hline 
DateOfDeath & 87 & 14 \\ \hline 
EmployeeOf & 75 & 16 \\ \hline 
FoundedBy & 326 & 30 \\ \hline 
FounderOf & 302 &  29 \\ \hline 
Titles & 26823 & 118 \\ \hline 
TopMembers & 314 & 26 \\ \hline

\end{tabular} 
\end{table}

% Note: Here is where we give as much stats as possible on our runs