### 2013-09-23 paper/kbanotebook2013/content/introduction.tex

\section{Introduction}

In this article we will describe the approach the GatorDSR team from the 
University of Florida in addressing the challenge at NIST’s Text 
REtrieval Conference (TREC) --- Knowledge Base Acceleration track 2013 
In the first section we describe 
the system design and articulate its evolution during the whole process, 
challenges and needs that we have overcome. In the second section we will go 
through the details of the NLP techniques we took at streaming slot filling
(SSF): given a slot name (entity + relation type) we fill in the slot value
(relation value) on structured text documents.

\ceg{Add a discussion of why machine driven kb population is important.}
The task is that given certain Wikipedia or Twitter entities 
(wiki/twitter URLs) and certain relations of interest given in Table 1, 
extract as many triple relations as possible (hence, slot filling). This can 
be used to automatically populate knowledgebases such as free-base or DBPedia 
or even filling in the information boxes at Wikipedia. The corpus is in 
English and is a snapshot of part of the web containing arXiv, social, news, 
linking, weblog, forum, classifieds, reviews, and memetracker documents. Each 
document is annotated using lingpipe and is called StreamItem, a bundle of 
StreamItems are put together and serialized as Apache Thrift objects, then 
compressed using xz compression with Lempel–Ziv–Markov chain algorithm (LZMA2) 
and finally encrypted using GNU Privacy Guard (GPG) with RSA asymmetric keys.
To have a sense of the scale of objects and compression as an example a 6mb
gpg.xz files would become 45 mb thrift objects which can contain a couple of 
thousand StreamItems depending on their size. Some of the documents have null 
values for their annotation fields.

Since this is \textit{streaming} slot filling, we are only interested in new 
slot values that were not substantiated earlier in the stream corpus, which 
ranges from October 2011 to February 2013. Below you can view the slot values 
and their types, the comprehensive description of which can be found at
\cite{tackbp} and \cite{aec}

\begin{table}[b]
\caption{Ontology of Slot Name Categories }
\centering

\begin{tabular}{|c|c|c|}
\hline 
PER & FAC & ORG \\ 
\hline 
\begin{tabular}{@{}l@{}}Affiliate \\ AssociateOf \\ Contact\_Meet\_PlaceTime \\ AwardsWon \\ 
DateOfDeath \\ 
CauseOfDeath\\
Titles\\
FounderOf\\
EmployeeOf\end{tabular}
  &
   \begin{tabular}[b]{l}Affiliate \\ Contact\_Meet\_Entity \end{tabular} 
   & 
   \begin{tabular}{@{}l@{}}Affiliate \\ TopMembers \\ FoundedBy\end{tabular} \\ 
\hline 
\end{tabular} 
\end{table}

Example 1: "Matthew DeLorenzo and Josiah Vega, both 14 years old and students 
at Elysian Charter School, were honored Friday morning by C-SPAN and received 
\$1,500 as well as an iPod Touch after winning a nationwide video contest."

target\_id:  http://en.wikipedia.org/wiki/Elysian\_Charter\_School

Affiliate:  "Matthew DeLorenzo"
Affiliate:  "Josiah Vega"

NOT an affiliate:  "C-SPAN"
NOT an affiliate:  "iPod Touch"

Example 2: "Veteran songwriters and performers Ben Mason, Jeff Severson and 
Jeff Smith will perform on Saturday, April 14 at 7:30 pm at Creative Cauldron 
at ArtSpace, 410 S. Maple Avenue."

target\_id: http://en.wikipedia.org/wiki/Jeff\_Severson

Affiliate: "Ben Mason"
Affiliate: "Jeff Severson"
Affiliate: "Jeff Smith"

NOT an Affiliate: "Creative Caldron"
NOT an Affiliate: "Art Space"

Example 3:  "Lt. Gov. Drew Wrigley and Robert Wefald, a retired North Dakota 
district judge and former state attorney general, unveiled the crest Friday 
during a ceremony at the North Dakota Capitol."

target\_id: http://en.wikipedia.org/wiki/Hjemkomst\_Center

Contact\_Meet\_PlaceTime: "Friday during a ceremony at the North Dakota Capitol"

In our approach we will perform pattern matching that conforms to each slot 
value along with post-processings to eliminate noisy outputs.

\ceg{How will each submission be evaluated?}

% Motivation

% KBA Task

% Evaluation Criteria

% Quick Discussion of our approach