### 2013-09-27 paper/kbanotebook2013/content/results.tex
Revision 2 on paper.

\section{Results}

%\ceg{We can add data on the total document size, The %number of each document type.  } 
 We had a single node cluster. The configuration of the server running the jobs is according to Table \ref{table:serverspec}. All together, The total size of the data after XZ compression and GPG encryption is 4.5TB and just over 500M StreamItems \cite{s3}. Data is stored in date-hour directories: From 2011-20-05-00 (12am) ) until 2013-02-13-23 (11pm) there are 11952 date-hour combinations. The source code of our system is stored as an open source project where enthusiasts can also contribute to \cite{github}, also the relevant discussion mailing list is accessible here \cite{googlegroups}.

 Our results are as follows: Document extraction using query entity matching with aliases, sentence 
extraction using alias matching and co-reference. Slot extraction using 
patterns, NER tags and NP tags. 158,052 documents with query entities, 17885 
unique extracted slot values for 8 slots and 139 entities, 4 slots and 31 
entities missing.

On the performance of the first submission run we performed some random 
sampling via two processes, the results of which are as in Table \ref{table:initialresult}. Our final submission has been elaborated in Tables \ref{table:finalresultaccuracy} and \ref{table:finalresultrecall}. You can observe that only CauseOfDeath is missing from the set of found slot values.

\begin{table}[b]
\caption{Benchmark Server Specifications }
\centering
\label{table:serverspec}
\begin{tabular}{|c|l|}
\hline 
Model & Dell xxx 32 cores \\ \hline 
OS & CentOS release 6.4 Final \\ \hline 
Software Stack & GCC version 4.4.7, Java 1.7.0\_25, Scala 2.9.2, SBT 0.12.3 \\ \hline 
 RAM & 64GB\\ \hline 
 Drives & 2x2.7TB disks, 6Gbps, 7200RPM\\ \hline 
\end{tabular} 
\end{table}

\begin{table}[t]
\caption{SSF Performance Measure on Initial Submission }
\centering
\label{table:initialresult}

\begin{tabular}{|c|c|c|c|}
\hline 
 & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Sampling \#1 & 55\% & 17\% & 27\% \\ 
\hline Sampling \#2 & 54\% & 15\% & 31\%  \\ 
\hline 
\end{tabular} 
\end{table}

\begin{table}[b]
\caption{SSF Accuracy Measure on Submission Infer }
\centering
\label{table:finalresultaccuracy}
\begin{tabular}{|c|c|c|c|}
\hline 
 & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Affiliate & 1\% & 95\% & 5\% \\ \hline 
AssociateOf & 63.6\% & 9.1\% & 27.3\%  \\ \hline 
AwardsWon & 10\% & 10\% & 80\%  \\ \hline 
CauseOfDeath & 0\% & 0\% & 0\%  \\ \hline 
Contact\_Meet\_Entity & 21\% & 42\% & 37\%  \\ \hline 
Contact\_Meet\_PlaceTime & 5\% & 20\% & 85\%  \\ \hline 
DateOfDeath & 29.6\% & 71\% & 25\%  \\ \hline 
EmployeeOf & 5\% & 30\% & 65\%  \\ \hline 
FoundedBy & 62\% & 17\% & 21\%  \\ \hline 
FounderOf & 50\% & 0\% & 50\%  \\ \hline 
Titles & 55\% & 0\% & 45\%  \\ \hline 
TopMembers & 33\% & 17\% & 50\%  \\ \hline

\end{tabular} 
\end{table}

\begin{table}[t]
\caption{Recall Measure on Submission Infer }
\centering
\label{table:finalresultrecall}
\begin{tabular}{|c|p{3cm}|p{4cm}|}
\hline 
 & Total instances of slot value found & \# of entities covered by slot value \\ 
\hline 
Affiliate & 108598 & 80 \\ \hline 
AssociateOf & 25278 & 106 \\ \hline 
AwardsWon & 38 & 14 \\ \hline 
CauseOfDeath & 0 & 0 \\ \hline 
Contact\_Meet\_Entity & 191 & 8 \\ \hline 
Contact\_Meet\_PlaceTime & 5974 & 109 \\ \hline 
DateOfDeath & 87 & 14 \\ \hline 
EmployeeOf & 75 & 16 \\ \hline 
FoundedBy & 326 & 30 \\ \hline 
FounderOf & 302 &  29 \\ \hline 
Titles & 26823 & 118 \\ \hline 
TopMembers & 314 & 26 \\ \hline

\end{tabular} 
\end{table}

% Note: Here is where we give as much stats as possible on our runs