### 2013-09-27 paper/kbanotebook2013/paper.tex

\documentclass[10pt]{article}

\usepackage{xcolor}
\usepackage{cite }
\usepackage{graphicx, epsfig}
\usepackage[margin=1in]{geometry}
\usepackage[hyphens]{url}

\pagenumbering{gobble} %Do not print page numbers

\newcommand\ceg[1]{\textcolor{red}{#1}}

\begin{document}

\title{GatorDSR: University of Florida Knowledge Base Acceleration Notebook}

%\numberofauthors{3}

\author {
  Morteza Shahriari Nia\footnote{ Department of Electrical \& Computer Engineering, University of Florida, Gainesville, Florida, USA}\\
       {m.shahriarinia@ufl.edu}
\and
  Christan Grant, Yang Peng, Daisy Zhe Wang\footnote{Department of Computer \& Information Science \& Engineering, University of Florida, Gainesville, Florida, USA}\\
       {\{cgrant, ypeng, daisyw\}@cise.ufl.edu}
\and
  Milenko Petrovic\footnote{Institute for Human and Machine Cognition, Ocala, Florida, USA}\\
       {mpetrovic@ihmc.us}
}

\maketitle

\input{content/abstract.tex}

\input{content/introduction.tex}

\input{content/system.tex}

\input{content/implementation.tex}

\input{content/results.tex}

\input{content/discussion.tex}

\input{content/conclusion.tex}

\section*{Acknowledgements}
Christan Grant is funded by a National Science Foundation Graduate Research
Fellowship under Grant No. DGE-0802270.

%\bibliographystyle{acm}
%\bibliography{citation}

\bibliographystyle{unsrt}
\bibliography{citation}

\end{document}