### 2013-10-02 paper/kbanotebook2013/content/results.tex

\section{Results}

%\ceg{We can add data on the total document size, The %number of each document type.  } 
 Our system was composed of a single node cluster, which we refer to as the \textit{server}; the configuration of which is as described in Table \ref{table:serverspec}. The corpus is a snapshot of the web in English. Each document is annotated using lingpipe and is called StreamItem, a bundle of 
StreamItems are put together and serialized as Apache Thrift objects, then 
compressed using xz compression with Lempel–Ziv–Markov chain algorithm (LZMA2) 
and finally encrypted using GNU Privacy Guard (GPG) with RSA asymmetric keys. The total size of the data after XZ compression and GPG encryption is 4.5TB and just over 500M StreamItems \cite{s3}. Data is stored in directories the naming of which is date-hour combination: from 2011-10-05-00 (5th of October 2011, 12am) until 2013-02-13-23 (13th of Feburary 2013, 11pm), which consists of 11952 date-hour combinations. This corpora consists of various media types the distribution of which can be found in \ref{table:documentsDist}. To have a sense of the scale of objects and compression as an example a 6mb
gpg.xz files would become 45 mb thrift objects which can contain a couple of 
thousand StreamItems depending on their size. Some of the documents have null 
values for their annotation fields. The first portion of the data which 
ranges from October 2011 to February 2013 is considered as training data. The source code of our system is stored as an open source project where enthusiasts can also contribute to \cite{github}, also the relevant discussion mailing list is accessible here \cite{googlegroups}.

\begin{table}[b]
\caption{Document Chunks Distribution }
\centering
\label{table:documentsDist}

\begin{tabular}{|c|l|}
\hline 
\textbf{\# of Documents} & \textbf{Document Type}\\ 
\hline 
10988 & 	arxiv (full text, abstracts in StreamItem.other\_content) \\ \hline
34887 & CLASSIFIED (spinn3r)  \\ \hline
77674 & FORUM (spinn3r)  \\ \hline
12947 & linking (reprocessed from kba-stream-corpus-2012, same stream\_id)  \\ \hline
141936 & MAINSTREAM\_NEWS (spinn3r)  \\ \hline
4137 & MEMETRACKER (spinn3r)  \\ \hline
280629 & news (reprocessed from kba-stream-corpus-2012, same stream\_id)  \\ \hline
6347 &  REVIEW (spinn3r) \\ \hline
688848 & 	 social (reprocessed from kba-stream-corpus-2012 plus extension, same stream\_id)  \\ \hline
740987 & WEBLOG (spinn3r)  \\ \hline

\hline 
\end{tabular} 
\end{table}

 Our results are as follows: Document extraction using query entity matching with aliases, sentence 
extraction using alias matching and co-reference. Slot extraction using 
patterns, NER tags and NP tags. 158,052 documents with query entities, 17885 
unique extracted slot values for 8 slots and 139 entities, 4 slots and 31 
entities missing.

On the performance of the first submission run we performed random 
sampling via two processes, the results of which are according to Table \ref{table:initialresult}. You can view that we have had an accuracy of around 55\%, and about 15\% wrong entity identified and 30\% incorrect value extracted across all entities and slot types. For our final submission, we provide a more detailed statistics, which has been elaborated in Tables \ref{table:finalresultrecall} and \ref{table:finalresultaccuracy}. Table \ref{table:finalresultrecall} shows the extent of search outreach for each slot name. You can see that \textit{Affiliate} has been the slot name with highest hits and \textit{CauseOfDeath} our lowest hit with 0 instances found matching our patterns, after that \textit{AwardsWon} has been the next with 38 instances found. Table \ref{table:finalresultaccuracy} addresses the relative accuracy measure per slot value. There you can view that we have had the highest accuracy of 63.6\% for \textit{AssociateOf} and the lowest of 1\% - 5\%  for \textit{Affiliate}, \textit{Contact\_Meet\_PlaceTime} and \textit{EmployeeOf}.

\begin{table}
\caption{Benchmark Server Specifications }
\centering
\label{table:serverspec}
\begin{tabular}{|c|l|}
\hline 
Model & Dell xxx 32 cores \\ \hline 
OS & CentOS release 6.4 Final \\ \hline 
Software Stack & GCC version 4.4.7, Java 1.7.0\_25, Scala 2.9.2, SBT 0.12.3 \\ \hline 
 RAM & 64GB\\ \hline 
 Drives & 2x2.7TB disks, 6Gbps, 7200RPM\\ \hline 
\end{tabular} 
\end{table}

\begin{table}
\caption{SSF Performance Measure on Initial Submission }
\centering
\label{table:initialresult}

\begin{tabular}{|c|c|c|c|}
\hline 
 & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Sampling \#1 & 55\% & 17\% & 27\% \\ 
\hline Sampling \#2 & 54\% & 15\% & 31\%  \\ 
\hline 
\end{tabular} 
\end{table}

\begin{table}
\caption{Recall Measure on Submission Infer }
\centering
\label{table:finalresultrecall}
\begin{tabular}{|c|p{3cm}|p{4cm}|}
\hline 
 Slot Name & Total instances of slot value found & \# of entities covered by slot value \\ 
\hline 
Affiliate & 108598 & 80 \\ \hline 
AssociateOf & 25278 & 106 \\ \hline 
AwardsWon & 38 & 14 \\ \hline 
CauseOfDeath & 0 & 0 \\ \hline 
Contact\_Meet\_Entity & 191 & 8 \\ \hline 
Contact\_Meet\_PlaceTime & 5974 & 109 \\ \hline 
DateOfDeath & 87 & 14 \\ \hline 
EmployeeOf & 75 & 16 \\ \hline 
FoundedBy & 326 & 30 \\ \hline 
FounderOf & 302 &  29 \\ \hline 
Titles & 26823 & 118 \\ \hline 
TopMembers & 314 & 26 \\ \hline

\end{tabular} 
\end{table}

\begin{table}
\caption{SSF Accuracy Measure on Submission Infer }
\centering
\label{table:finalresultaccuracy}
\begin{tabular}{|c|c|c|c|}
\hline 
 Slot Name  & Correct & Incorrect Entity name & Incorrect Value \\ 
\hline 
Affiliate & 1\% & 95\% & 5\% \\ \hline 
AssociateOf & 63.6\% & 9.1\% & 27.3\%  \\ \hline 
AwardsWon & 10\% & 10\% & 80\%  \\ \hline 
CauseOfDeath & 0\% & 0\% & 0\%  \\ \hline 
Contact\_Meet\_Entity & 21\% & 42\% & 37\%  \\ \hline 
Contact\_Meet\_PlaceTime & 5\% & 20\% & 85\%  \\ \hline 
DateOfDeath & 29.6\% & 71\% & 25\%  \\ \hline 
EmployeeOf & 5\% & 30\% & 65\%  \\ \hline 
FoundedBy & 62\% & 17\% & 21\%  \\ \hline 
FounderOf & 50\% & 0\% & 50\%  \\ \hline 
Titles & 55\% & 0\% & 45\%  \\ \hline 
TopMembers & 33\% & 17\% & 50\%  \\ \hline

\end{tabular} 
\end{table}

% Note: Here is where we give as much stats as possible on our runs