### 2013-10-11 paper/kbanotebook2013/content/introduction.tex

\section{Introduction}

An important challenge in maintaining Wikipedia.org (WP), the most popular 
web-based, collaborative, multilingual knowledge base on the internet, is  
making sure its contents are up-to-date. Presently, there is considerable time lag 
between the publication date of cited news and the date of an edit to WP 
creating the citation. The median time lag for a sample of about 60K
web pages cited by WP articles in the \textit{living\_people} category is over 
a year and the distribution has a long and heavy tail~\cite{JFrank12}. 
Also, the majority of WP entities have updates on their associated article much 
less frequently than their mention frequency. Such stale entries 
are the norm in any large knowledge base (KB) because the number of humans 
maintaining the KB is far fewer than the number of entities. 
%Further, the number of mentions is much larger than the number of entities~\cite{JFrank12}. 
Reducing latency keeps KBs such as WP relevant and helpful it users.

Given an entity page, possible citations may come from a variety of sources.
The actual citable information is a small percentage of the total documents that appear on the web.
We develop a system to read streaming data and filter out articles that are candidates for citations.
Given these documents, we extract the pairs of information in the 
article that is recommended for citations for each knowledge base page.

Our system contains three main components. First, we pre-process the data and
build models representing the knowledge base entries.
Next, we use the models to filter a stream of documents so they only contain 
candidate recommendations.
Lastly, we processes sentences from candidate extractions and return 
specific information to be cited.

We built this system as part of a formal task described in Section~\ref{sec:kbatask}.
In this paper, we describe the system we built to process the data.
Our approach to this task is to build a modular system
that allows us to explore the nuances of the training data and queries.
Overall, our contributions are the following:
\begin{itemize}[noitemsep,nolistsep]
\item we introduce a method to build models of name variations (Section~\ref{section:aliasgeneration});
\item we built a system to filter a large amount of diverse documents (Section~\ref{sec:ccrimpl});
\item we extract entity-slot-value triples of information to be added to KB (Section~\ref{sec:ssve});
\item we filter the final results using deduplication and inference (Section~\ref{section:highAccuracyFilter});
\item we self-evaluate our results over a 4.5 TB data set (Section~\ref{sec:results}).
\end{itemize}

% Evaluation Criteria

% Quick Discussion of our approach