### 2013-10-25 paper/flairs27/content/abstract.tex
added the paper for FLAIRS27. It needs a LOT of work

\begin{abstract}

% TODO this can be updated at the end

In this paper we will present the system design and algorithm adopted by the 
GatorDSR team, University of Florida to efficiently process TREC KBA 2013 --- 
SSF track. Here we will describe the system as well as the details 
the algorithms used to extract slot values for the given slot name. Scalability, efficiency, precision and recall are the major goals of this work, 
given the overly limited time limitation and available computational resources.

\end{abstract}