### 2013-11-03 paper/flairs27/content/abstract.tex

\begin{abstract}

% TODO this can be updated at the end

In this paper we address extracting central slot values from internet pertinent to entities in wikipedia, which is the most popular web-based, colaborative multilingual encyclopieda on the internet. This is comparable to Freebase in a query-driven manner. Our contributions are two fold, first we design a system to efficiently find central documents on the internet that contain directly citable information regarding a given wikipedia entity; second, we design a system to extract specific features of that entity from central documents which is refered to as slot filling. Our results demonstrate that the system is very efficient in the web scale nature of the problem: being highly memory and I/O efficient and that we can achieve high accuracy and recall for given slot values.

\end{abstract}