### 2013-11-04 paper/flairs27/content/abstract.tex

\begin{abstract}

% TODO this can be updated at the end

In this paper we address extracting slot values from internet pertinent to entities in wikipedia, which is the most popular web-based, colaborative multilingual encyclopieda on the internet. This is comparable to Freebase in a query-driven manner. Our contributions are two fold, first we design a system to efficiently find central documents on the internet that contain directly citable content regarding a given wikipedia entity; second, we design a system to extract certain slot-values of the entity from those documents. Our results demonstrate that the system is very efficient in the web scale nature of the problem: being highly memory and I/O efficient and that we can achieve high accuracy and recall for given slot values.

\end{abstract}