### 2013-11-04 paper/flairs27/content/appendix.tex

\appendix

\section{KBA Task Background}
\label{sec:kbatask}
The National Institute of Standards (NIST) hosted the Text REtrieval Conference (TREC) --- Knowledge Base Acceleration challenge in 2013. The task contains two main sections designed for this track, Cumulative Citation Recommendation and Streaming Slot Filling. Due to the importance of knowledge bases, both of these 
tracks aim to accelerating populating them, hence the title Knowledge Base Acceleration (KBA). Below we describe each of these tracks and their purposes.

\subsection{Cumulative Citation Recommendation (CCR)}

For this track, assessors were instructed to ``use the Wikipedia article to identify (disambiguate) the entity, and then imagine forgetting all info in the WP article and asking whether the text provides any information about the entity'' \cite{JFrank12}. Documents are divided according if an entity is mentioned and a relevance level to the entity.

More specifically, a document may have a mention or be without a mention.
\begin{itemize}[noitemsep]
  \item \textbf{Mention:} Document explicitly mentions target entity, such as full 
    name, partial name, nickname, pseudonym, title, stage name.
  \item \textbf{Zero-mention:} Document does not directly mention target. Could 
    still be relevant, e.g.\ metonymic references like ``this administration'' 
    $\rightarrow$ ``Obama''. See also synecdoche. A document could also be relevant to 
    target entity through relation to entities mentioned in document -- apply this 
    test question: can I learn something from this document about target entity using 
    whatever other information I have about entity?
\end{itemize}

The relevance of a document to a query is split into the four classifications.
\begin{itemize}[noitemsep]
  \item \textbf{Garbage:} not relevant, e.g.\ spam.
  \item \textbf{Neutral:} Not relevant, i.e.\ no info could be deduced about entity, 
    e.g., entity name used in product name, or only pertains to community of target 
    such that no information could be learned about entity, although you can see how 
    an automatic algorithm might have thought it was relevant.
  \item \textbf{Relevant:} Relates indirectly, e.g., tangential with substantive 
    implications, or topics or events of likely
    impact on entity.
  \item \textbf{Central:} Relates directly to target such that you would cite it in 
    the WP article for this entity, e.g.\ entity is a
    central figure in topics/events.
\end{itemize}

\subsection{Streaming Slot Filling (SSF)}
The task is that given certain WP or Twitter entities (wiki/twitter URLs) and certain relations of interest (given in Table~\ref{table:slotNameOntology}), extract as many triple relations as possible (hence, slot filling). This can be used to automatically populate knowledgebases such as free-base or DBPedia or even fill-in the information boxes at Wikipedia. Below, you can view some examples of what it means to fill in a slot value; in each example there is a sentence of interest that we wish to extract slot values from, an entity that the slot value is related to, and a slot name which can be thought of as the topic of the slot value:

\noindent \textbf{Example 1:} ``Matthew DeLorenzo and Josiah Vega, both 14 years old and students at Elysian Charter School, were honored Friday morning by C-SPAN and received \$1,500 as well as an iPod Touch after winning a nationwide video contest.''

Entity:  http://en.wikipedia.org/wiki/Elysian\_Charter\_School

Slot name: Affiliate

Possible slot values: ``Matthew DeLorenzo'', ``Josiah Vega''

Incorrect slot values: ``C-SPAN'', ``iPod Touch''

\noindent \textbf{Example 2:} ``Veteran songwriters and performers Ben Mason, Jeff Severson and Jeff Smith will perform on Saturday, April 14 at 7:30 pm at Creative Cauldron at ArtSpace, 410 S. Maple Avenue.''

Entity: http://en.wikipedia.org/wiki/Jeff\_Severson

Slot name: Affiliate

Possible slot values: ``Ben Mason'', ``Jeff Severson'', ``Jeff Smith''

Incorrect slot values: ``Creative Caldron'', ``Art Space''

\noindent \textbf{Example 3:}  ``Lt. Gov. Drew Wrigley and Robert Wefald, a retired North Dakota district judge and former state attorney general, unveiled the crest Friday during a ceremony at the North Dakota Capitol.''

Entity: http://en.wikipedia.org/wiki/Hjemkomst\_Center

Slot name: Contact\_Meet\_PlaceTime

Slot value: ``Friday during a ceremony at the North Dakota Capitol''

\noindent In \textit{streaming} slot filling, we are only interested in new slot values that were not substantiated earlier in the stream corpus. In Table~\ref{table:slotNameOntology} you can view some slot values and their types, the comprehensive description of which can be found at \cite{tackbp} and \cite{aec}. The details of the metric for SSF will favor systems that most closely match the changes in the ground truth time line of slot values. This is done by searching for other documents that mentioned an entity and exactly 
matched the slot fill strings selected by the assessors.

%In the rest of the paper, we address the following material. In Section 2, we 
%sketch the main components of the system and their purposes. In Section 3, 
%describe the details of how we designed and implemented certain components 
%critical to the behavior of the system. Section 4 addresses the performance and 
%the results we achieved. Section 5, goes through a discussion of the challenges we 
%had to produce reliable results over the massive amount of information available, 
%and how we overcome these issues.
For similar approaches regarding last year's track you can refer to \cite{ji2011knowledge}.

%\ceg{Add a discussion of why machine driven kb population is important.}

%\ceg{How will each submission be evaluated?}

% Motivation

% KBA Task