### 2013-11-04 paper/flairs27/content/implementation.tex

\section{Implementation}
\label{section:implementation}

We extract aliases for entities from Wikipedia automatically both using API and using the actual page content, then apply pattern matching rules for slot value extraction. Our contribution is that we perform pattern matching that conforms to each slot value along with post-processings to eliminate noisy outputs.

%\ceg{Instead of this paragraph we talk about what this %section will
%contain. We can discuss the input/output of the system %too.
%}

\subsection{Alias Generation}
\label{section:aliasgeneration}

\subsection{Cumulative Citation Recommendation}
\label{sec:ccrimpl}

\subsection{Streaming Slot Value Extraction}

% Note: Talk about how each of the algorithms were created

% Note: Give enough information for our algorithms to be
% reimplemented and verified.