### 2013-11-06 paper/flairs27/content/discussion.tex

\section{Discussions}

%In this section we address some of the challenges that we faced during the course of this project. We started 
%off by trying to use the off-the-shelf tools for this task. We started 
%by using Scala/Spark~\cite{ferc11} to benefit from the parallelization 
%performance there. Unfortunately Spark was not performant and the distributed reliability overhead 
%was much more than tolerable. We moved on to use Scala parallelization  framework itself, and unfortunately it did not satisfy our needs as well; there was 
%excessive memory overhead on map-reduce jobs. We migrated the core of the 
%system to Java Parallelism APIs and that was not good enough either, we still demanded 
%better. So we built our own parallel system which in actual performance had 
%the least of overhead, the least memory consumption and the most robust memory 
%model which avoided unpredictable CPU stalls on garbage collections in 
%processing the corpus.

Table~\ref{table:finalresultrecall} show a varied distribution of extracted slot names. Some slots naturally have more results than other slots. For example, AssociateOf and Affiliate have more slot values than DateOfDeath and CauseOfDeath, since there are only so few entities that are deceased. Also, some patterns are more general causing more extractions. For example, for Affiliate, we use \textit{and}, \textit{with} as anchor words. These words are more common than \textit{dead} or \textit{died} or \textit{founded} in other patterns.

As part of future work regarding enhancing precision, we can focus on fixing the wrong entities found, remove noisy tags by the Lingpipe or use more accurate NLP taggers such as Stanford NLP (Due to the fact that Stanford NLP is very slow we avoided using it, as due to speed its use was out of question) and finally use better patterns to enahnce results matched by the patterns.  On the other hand to increase recall we need to find better aliases and go for more resources to discover other ways that an entity might be addressed (e.g. twitter id, website, etc), add powerful patterns that can capture more cases of slot values. We would also use entity resolution methods and other advanced methods to improve 
the accuracy and recall of entity extraction part.

For slot extraction, to improve the performance, we need: 1) Using multi-class classifiers instead of pattern matching method to extract slot values in order to increase both recall and accuracy for slots ``Affiliate'', ``AssociateOf'', ``FounderOf'', ``EmployeeOf'', ``FoundedBy'', ``TopMembers'', ``Contact\_Meet\_Entity'' and so on. 2) For special slots, like ``Titles'', ``DateOfDeath'', ``CauseOfDeath'', ``AwardsWon'', using different kind of advanced methods, e.g.\ classifiers, matching methods. 3) Using other NLP tools or using classifiers to overcome the drawbacks of the LingPipe’s inaccurate tags. The first and second tasks are the most important tasks we need to do.

Some of the entities are not popular. For example, a `Brenda Weiler' Google search result has 860,000 documents over the whole web. For our small portion of the web it might make sense. The histogram of the entities shows that more than half of the entities have appeared in less than 10 StreamItems. A good portion have appeared only once.

%A theory is that we are falling behind because we are using the cleansed (from HTML tags)
%version of the corpus. We believe there has been a reason that TREC has 
%included the actual HTML document as well as the cleansed version. This 
%definitely will convey some information to us. If it was as easy as reading 
%some clean text they wouldn't bother including so much data for teams to be 
%useless. So we guess is that we are missing some information from not using 
%the actual document. And, we are looking for tokens with entity value set 
%which will depend us dramatically on the accuracy of lingpipe, which is a fast 
%algorithm but is not as good as other NLP tools can be e.g. Stanford NLP.

% Note: Here we discuss why the results are the way they are
% Give pros and cons, talk about how the implemented algorithms
% performed in the actual implementation.
% Answer the `why' question about all the trends in the results section.

Alltogether, We experimented through different tools and approaches to best process the massive amounts of data on the platform that we had available to us. We generate aliases for wikipedia entities using Wiki API and extract some aliases from wikipedia pages text itself. We process documents that mention entities for slot value extraction. Slot values are determined using pattern matching over coreferences of entities in sentences. Finally post processing will filter, cleanup and infers some new slot values to enhance recall and accuracy.

We noticed that some tools that claim to be performant for using the hardware capabilities at hand sometimes don't really work as claimed and you should not always rely on one without a thorough A/B testing of performance which we ended up in generating our in-house system for processing the corpus and generating the index. Furthermore, on extracting slot values, pattern matching might not be the best options but definitely can produce some good results at hand. We have plans on generating classifiers for slot value extraction purposes. Entity resolution on the other hand was a topic we spent sometime on but could not get to stable grounds for it. Entity resolution will distinguish between entities of the same name but different contexts. Further improvements on this component of the system are required.

We sampled documents from the training data period to generate an initial set of patterns. We then use these patterns to generate  results. By manully looking at these results, we prune some patterns with poor ­performance and add more patterns that we identified from these results. We use several iterations to find the best patterns. We found that it is very time consuming to identity quality patterns.