### 2013-11-18 paper/flairs27/paper.tex
Ready for submission

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{xcolor}
\usepackage{cite }
\usepackage{graphicx, epsfig}
\usepackage[margin=1in]{geometry}
\usepackage[hyphens]{url}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algorithmic} % noend will ksip printing end of blocks

\usepackage{cite}
\usepackage{natbib}

\usepackage{balance}

\pagenumbering{gobble} %Do not print page numbers

\newcommand\ceg[1]{\textcolor{red}{#1 -- CEG}} % Red marker to declare errors.
\newcommand\hide[2]{#2}

\begin{document}

\title{Streaming Fact Extraction for Wikipedia Entities at Web-Scale}

%\numberofauthors{3}

\author{Morteza Shahriari Nia, Christan Grant, Yang Peng, Daisy Zhe Wang\footnote{University of Florida, Gainesville, Florida, USA}, Milenko Petrovic\footnote{Institute for Human and Machine Cognition, Ocala, Florida, USA}\\
       \\
       {\{msnia, cgrant, ypeng, daisyw\}@cise.ufl.edu,}
       {mpetrovic@ihmc.us}
}
%\author{XXX}

\maketitle

\input{content/abstract.tex}

\input{content/introduction.tex}

\input{content/system.tex}

%\input{content/implementation.tex}

\input{content/results.tex}

\input{content/discussion.tex}

\input{content/conclusion.tex}

\section*{Acknowledgements}
%We would like to thank \hide{John Frank}{XXX} and \hide{NIST}{XXX} for providing the data set.
%\hide{Christan Grant}{XXXX} is funded by \hide{a National Science Foundation Graduate Research Fellowship under Grant No. DGE-0802270}{XXXX}.

Christan Grant is funded by a National Science Foundation Graduate Research Fellowship under Grant No. DGE-0802270.

%\bibliographystyle{acm}
%\bibliography{citation}

\bibliography{citation}
\bibliographystyle{aaai}

%\input{content/appendix.tex}

\end{document}