### 2013-11-18 paper/flairs27/content/abstract.tex
updated abstract. Re-writted, most jargon removed.

\begin{abstract}

% TODO this can be updated at the end

%In this paper we address extracting slot values from internet pertinent to entities in wikipedia, which is the most popular web-based, colaborative multilingual encyclopieda on the internet. This is comparable to Freebase in a query-driven manner. Our contributions are two fold, first we design a system to efficiently find central documents on the internet that contain directly citable content regarding a given wikipedia entity; second, we design a system to extract certain slot-values of the entity from those documents. Our results demonstrate that the system is very efficient in the web scale nature of the problem: being highly memory and I/O efficient and that we can achieve high accuracy and recall for given slot values.

Wikipedia.org is the largest online resource for free information and maintained by a small number of volunteer editors.
Because of its large size, information on some pages can easily be neglected and can easily become out of date.
To address the issues of stale articles we create a system that is able to read
in a stream in a web of diverse documents from across the web and recommend
facts to be added to specified Wikipedia pages.
We developed a three-stage stream system to create models of entities,
filter out irrelevant documents and 
extract facts from documents mentioning entities.
The systems over a 4.5 TB web document corpus data set over 16 months and 139 Wikipedia pages.
Our results show a promising framework for fact extraction from arbitrary web pages for Wikipedia pages.

\end{abstract}